Refunctionalization of Abstracting Abstract Machines (Functional Pearl)

### 2018-03-17 paper.tex
update intro

Abstracting abstract machines (AAMs) and abstract definitional interpreters (ADIs) are, broadly speaking, both examples of abstract interpreters. \note{JD: need more bullet points here and I can fill this in}

@@

These transformations are used throughout the paper, with refunctionalization and
defunctionalization of abstract interpreters playing important roles for the call stack of
the analyzed language (i.e., Scala)\note{right?}. All of this insight allows the revelation of
precisely \emph{why} small-step AAM does not have the pushdown control-flow property
(without extra effort), as well as explains why ADI naturally extends pushdown control-flow
property from its defining language.

@@

% \item A new understanding of pushdown control-flow analysis by
%   refunctionalization.
%   We present an abstract interpreter obtained from refunctionalizing
%   a small-step abstracting abstract machine.
%   This abstract interpreter naturally exhibits the pushdown control-flow property,
%   as the refunctionalization forces us to sequentialize the
%   abstract execution.
%   We also analyze its correctness, soundness, and termination.

@@

\subsection{Abstracting Abstract Machines} \label{aam}
Abstracting abstract machines (AAM) is a systematic methodology that derives sound
abstract interpreters for higher-order functional languages from concrete
abstract machines \cite{van2012systematic, van2010abstracting}.
An abstracting abstract machine implements computable abstract semantics that
approximates the runtime behaviors of programs.
Since the state space of concrete execution is possibly infinite,
to analyze programs, the key insight of AAM approach is to allocate both bindings
and continuations on the store, and then bound the addresses space to be finite.
Since each component of state is finite,
the abstracted machine-state space is finite, and computable.

In this section, we derive the abstracting abstract machine from concrete
CESK machines, and also show how to instantiate
useful $k$-call-sensitive control-flow analysis.

@@

\textbf{The Problem of Return-flows.}
Pushdown control-flow is a property in analysis that precisely models
the runtime call-stack of the analyzed program.
A pushdown control-flow analysis provides as exact as runtime return-flow
when analyzing the program, but traditional control-flow analysis collapses
the state space into a finite space thus causes imprecise stack modeling.

To see how traditional control-flow analysis suffers from spurious return-flows,
we can consider the following example:

@@

For deterministic languages, a reduction semantics is a structural operational
semantics in continuation style, where the reduction context is a
defunctionalized continuation.
Unfortunately, the abstract abstract machines is non-deterministic, where
the non-determinism is handled by a work list.
We can turn this explicit non-determinism into the reduction context of
AAM.

\textbf{Pushdown Control-flow Analysis.}
In recent years, there are significant efforts \cite{vardoulakis2010cfa2, earl2012introspective,
gilray2016pushdown, johnson2015abstracting} to achieve precise call/return
match based on small-step abstracting abstract machines.

>>>>

Abstracting abstract machines (AAM) is a systematic methodology proposed by \citeauthor{van2012systematic}
for constructing abstract interpreters of higher-order functional languages \cite{van2012systematic, van2010abstracting}. 
Given a concrete small-step abstract machine, (e.g, the CESK machine, Krivine's machine or etc), 
by allocating continuations in the store and bound both the value addresses and continuation
addresses to be finite, we obtain an abstract interpreter with finite state space which 
can be used for performing sound static analysis.
One can further instantiate different polyvariant control-flow analysis by using different address allocators \cite{Gilray:2016:ACP:2951913.2951936}.

Applying the same idea on big-step definitional interpreters, \citeauthor{darais2017abstracting} 
build abstract definitional interpreters (ADI) that are written in monadic style \cite{darais2017abstracting}
By changing the underlying
monads, the definition of interpreter has no change, but we can recover different semantics 
including the concrete semantics and various abstract semantics.

Broadly speaking, abstract abstract machines and abstract definitional interpreters are both
abstract interpreters. They are obtained by applying a combination of abstractions to their
concrete counterparts, abstract machines and definitional interperters, perspectively.
The relations between abstract machines, definitional interpreters, and monadic interpreters 
is intensively studied by Danvy and his collaborator in the past years 
\cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268}.

But in terms of their abstract semantic artifacts, 
the relations between small-step abstract abstract machines and big-step abstract 
definitional interpreters, and the question that how can we interdrive each other is not
well studied.
In this functional peral, we practical answer this question by presenting a series of 
syntactical transformations on the program from small-step abstract abstract machines 
to big-step abstract definitional interpreters.

\subsection{outline}

@@

These transformations are used throughout the paper, with refunctionalization and
defunctionalization of abstract interpreters playing important roles for the call stack of
the analyzed language. All of this insight allows the revelation of
precisely \emph{why} small-step AAM does not have the pushdown control-flow property
(without extra effort), as well as explains why ADI naturally extends pushdown control-flow
property from its defining language.

@@

% \item A new understanding of pushdown control flow analysis by
%   refunctionalization.
%   We present an abstract interpreter obtained from refunctionalizing
%   a small-step abstracting abstract machine.
%   This abstract interpreter naturally exhibits the pushdown control-flow property,
%   as the refunctionalization forces us to sequentialize the
%   abstract execution.
%   We also analyze its correctness, soundness, and termination.

@@

\subsection{Abstracting Abstract Machines} \label{aam}
Abstracting abstract machines (AAM) is a systematic methodology that derives sound
abstract interpreters for higher-order functional languages from concrete
abstract machines \cite{van2012systematic, van2010abstracting}.
An abstracting abstract machine implements computable abstract semantics which
approximates the runtime behaviors of programs.
Since the state space of concrete execution is possibly infinite,
the key insight of AAM approach when analyzing programs is to allocate both bindings
and continuations on the store, and then bound the addresses space to be finite.
Since each component of state is finite,
the abstracted machine-state space is also finite, and therefore computable.

In this section, we derive the abstracting abstract machine from concrete
CESK machines, and also show how to instantiate
useful $k$-call-sensitive control flow analysis.

@@

\textbf{The Problem of Return-flows.}
Pushdown control-flow is a property in analysis that precisely models
the runtime call-stack of the analyzed program.
A pushdown control flow analysis provides as exact as runtime return-flow
when analyzing the program, but traditional control flow analysis collapses
the state space into a finite space thus causes imprecise stack modeling.

To see how traditional control flow analysis suffers from spurious return-flows,
we can consider the following example:

@@

For deterministic languages, a reduction semantics is a structural operational
semantics in continuation style, where the reduction context is a
defunctionalized continuation.
Unfortunately, the abstract abstract machines is nondeterministic, where
the nondeterminism is handled by a work list.
We can turn this explicit nondeterminism into the reduction context of
AAM.

\textbf{Pushdown Control flow Analysis.}
In recent years, there are significant efforts \cite{vardoulakis2010cfa2, earl2012introspective,
gilray2016pushdown, johnson2015abstracting} to achieve precise call/return
match based on small-step abstracting abstract machines.
